### 2022-02-18 paper1.Rmd

---
title: "Article 1:	Construct Validity and Correspondence of Google Trends"
author: "Kelsey Gonzalez"
date: "`r format(Sys.time(), '%d %B, %Y')`"
editor_options:
  chunk_output_type: console
output:
  bookdown::pdf_document2:
    latex_engine: xelatex
    number_sections: true
  bookdown::word_document2: 
    toc: true
    reference_docx: StylesTemplate.docx
header-includes:
   - \usepackage{siunitx}
   - \usepackage{setspace}
   - \usepackage{dcolumn}
   - \usepackage{caption}
   - \usepackage{longtable}
   - \usepackage{booktabs}
   - \doublespacing
   - \usepackage{placeins}
csl: american-sociological-association.csl
bibliography: "thesis.bib"
indent: true
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      message = FALSE,
                      warning = FALSE,
                      # cache = TRUE,
                      fig.width = 7,
                      fig.asp = 0.8,
                      out.width = "80%",
                      fig.align="center"
                      # dev.args = list(png = list(type = "cairo"))
                      )

if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, glue, extrafont, here, jtools, psych,
               modelsummary, showtext, thematic, kableExtra, scales)

my_palette = c('#56b4e9','#999999','#E69F00','#009E73','#CC79A7','#0072B2','#D55E00','#E69F00')
my_palette  <- c(MetBrewer::met.brewer(name = 'Cross', type = 'discrete'), '#555F61')

## Automatically use showtext to render text
font_add_google("Lora", "lora")
showtext_auto()

theme_diss <- function(base_size = 14) {
  theme_minimal(base_size = base_size) %+replace%
    theme(
      # Figure assembly
      plot.title = element_text(family="lora",
                                size = rel(1), 
                                margin = margin(0,0,5,0), 
                                hjust = 0),
      plot.title.position = "plot",
      plot.subtitle = element_text(family="lora",
                                   size = rel(0.85)),
      plot.caption = element_text(family="lora",
                                  size = rel(0.70),
                                  hjust = 1),
      # Graphical Zones
      panel.grid.minor = element_blank(),
      panel.border = element_blank(),
      # Axes
      axis.title = element_text(family="lora", 
                                size = rel(0.85)),
      axis.text = element_text(family="lora", 
                               size = rel(0.70)),
      axis.line = element_line(color = "black", 
                               arrow = arrow(length = unit(0.3, "lines"), 
                                             type = "closed")),
      # Legend
      legend.title = element_text(family="lora",
                                  size = rel(0.85)),
      legend.text = element_text(family="lora", 
                                 size = rel(0.70)),
      legend.key = element_rect(fill = "transparent", 
                                colour = NA),
      legend.key.size = unit(1.5, "lines"),
      legend.background = element_rect(fill = "transparent", 
                                       colour = NA),
      # Facetting 
      strip.background = element_rect(fill = "#17252D",
                                      color = "#17252D"),
      strip.text = element_text(family="lora",
                                size = rel(0.85), 
                                color = "white", 
                                margin = margin(5,0,5,0))
    )
}

theme_set(theme_diss())

conf_int_prop <- function(var){
  p <- mean(!!var, na.rm = T)
  q <- 1 - p
  n <- nrow(rain)
  min <- p - 1.96*sqrt((p*q)/n)
  max <- p + 1.96*sqrt((p*q)/n) 
  paste0(label_percent()(min), ", ", label_percent()(max))
}
print_perc <- function(var){
  label_percent()(mean(var, na.rm = T))
}

```

# Abstract

Possible titles 
Digital Trace Data as Social Indicators: Indicators of Attention Rather Than Support

**Keywords**:

# Intro 
New big data sources have led to vast possibilities for social science research
because they are bigger, cheaper, and already available
[@kingEnsuringDataRichFuture2011; @lazerComputationalSocialScience2009;
@salganikBitBitSocial2017]. Before overenthusiastically embracing these sources
into our workflows, social scientists must clearly establish parameters under
which these data sources could be operationalized
[@bailCulturalEnvironmentMeasuring2014; @lazerParableGoogleFlu2014]. As prior
research outlined, the “quantity of data does not mean that one can ignore
foundational issues of measurement and construct validity and reliability and
dependencies among data” [@lazerParableGoogleFlu2014, p. 1203]. Building on this
prior research outlining these various issues with big data
[@boydCriticalQuestionsBig2012; @lazerIssuesConstructValidity2015], this paper
tests the construct validity of Google Search Trends as an indicator of three
different cases, namely cultural attitudes, disease prevalence and voting
behavior. These three cases will be tested using cultural indicators from the
NORC General Social Survey, United States county-level suicide rates from
National Center for Health Statistics, US cases rates of Covid-19 from The New
York Times, historical US Presidential Election results, and the American
National Election Survey. Data will be analyzed against corresponding
operationalized Google Trends longitudinal data using Pearson’s r for pairwise
correlations, testing for the strength of relationships between the Google Trend
indicator and the respective comparison indicator. This paper will contribute to
the creation of methodological norms and standards of how to use Google Trends
as a big data source for societal research and serve as a critical inquiry into
the adoption of big data without a critical eye for the ecological validity of
the sources.

With the expansion of big data, some research has shown extremely innovative
methods that lead to groundbreaking results that are shown to be reliable. As an
example, @blumenstockPredictingPovertyWealth2015 use
county-level cell-phone records to construct the distribution of poverty and
wealth in Rwanda, a country where national surveys and censes are rare and
costly. However, @blumenstockPredictingPovertyWealth2015
go to great lengths to demonstrate that their operationalization of the cell
phone data creates a reliable and valid construct; few social science papers
utilizing big data dig into the construct validity of their metrics to this
extent and even fewer publications focus on methodological guidelines of how to
use sources of big data [@asseoTrackingCOVID19Using2020;
@stilesAssessingCriterionValidity2018]. However, research has shown the small
adjustments to an algorithm or metric may void any research insight we are able
to pull from such data [@lazerParableGoogleFlu2014]. Because of this, I propose
a methodological validation study of the Google Search Trends source of big data
to investigate how it is advisable to utilize this data in social scientific
research.

I will use three categorizations of ways I propose Google Trends could be
operationalized for social scientific usage. First, I’ll test Google Trends as
an operationalization of cultural attitudes with the General Social Survey.
After @bailCulturalEnvironmentMeasuring2014’s call for cultural
sociologists to utilize the ever-expanding world of big data, Google trends as a
data source began appearing in sociological and social science research.  From
research on mass shootings and firearms [@brownsteinInternetSearchPatterns2020;
@semenzaInformationseekingWakeTragedy2020], protest and anti-Muslim sentiment
[@bailUsingInternetSearch2018; @barrieSearchingRacismGeorge2020;
@grossThereFergusonEffect2017], to analyzing country-level changes in social
perception [@reyes_etal18], Google search trends are a new and innovative
indicator of cultural interest. Extending into social networks and culture,
@bailPrestigeProximityPrejudice2019 even used Google
trends to measure how culture spreads around the globe.

Google Search Trends have also been used continuously in estimations of disease
prevalence and population health in journals like the Journal of Medical
Internet Research. While much of this research has focused on the Covid-19
pandemic [@jimenez_etal20; @jimenezCOVID19SymptomGoogle2020;
@limEstimatingInformationSeekingBehaviour2020;
@mavraganiCOVID19PredictabilityUnited2020; @nguyenGoogleTrendsAnalysis2020;
@todorovaInternetBasedData2021] , other research has investigated Google Trends
as an indicator of wellbeing [@brodeurCOVID19LockdownsWellbeing2021;
@carpiTwitterSubjectiveWellBeing2020; @duCOVID19IncreasesOnline2020],
suicidality [@burnettTimeTrendsPublic2020], vaccination uptake
[@dalumhansenEnsembleLearnedVaccination2016], obesity
[@sarigulNowcastingObesityUsing2014], and even insomnia
[@zittingGoogleTrendsReveal2020], to cover a few examples. For a partial
review of other utilizations, see @nutiUseGoogleTrends2014. According to
@jaidkaInformationseekingVsSharing2021, the majority of studies profess a
correlation of > .70, “demonstrating the vast potential of Google Search as a
proxy for monitoring population health” (p. 3) based on assumptions that
individuals search because of self-diagnosis and to identify possible courses
of treatment [@dechoudhurySeekingSharingHealth2014].

<!-- Following reports say that Google searches can predict covid  -->
<!-- 14. Brunori, P. & Resce, G. Searching for the peak Google Trends and the Covid-19 outbreak in Italy. SERIES working papers N. 04/2020 -->
<!-- Lampos, V. et al. Tracking COVID-19 using online search. Preprint at https://arxiv.org/abs/2003.08086 (2020). -->
<!-- 16. Walker, A., Hopkins, C. & Surda, P. The use of google trends to investigate the loss of smell related searches during COVID‐19 -->
<!-- outbreak. Int. Forum Allergy Rh. 10, 839–847 (2020). -->
<!-- 17. Stephens-Davidowitz, S. Google Searches Can Help Us Find Emerging Covid-19 Outbreaks. The New York Times https://www. -->
<!-- nytimes.com/2020/04/05/opinion/coronavirus-google-searches.html (2020). -->
<!-- 18. Frank, T. Goldman says fewer ‘loss of smell’ Google queries suggest better COVID outlook. Consumer News and Business Channel. CNBC https://www.cnbc.com/2020/04/15/goldman-says-fewer-loss-of-smell-google-searches-is-a-positive-sign-for-pande -->
<!-- mic.html (2020) -->

Various sources have also used Google Trends as a way to forecast political
elections and political attitudes [@wolfTrendingRightDirection2018]. For
instance, @swearingenGoogleInsightsSenate2014 investigate how U.S. Senate
Elections relate to attention measured by search traffic.
@prado-romanGoogleTrendsPredictor2020 compare how Google Search trends are able
to predict presidential election results in both the United States and Canada.
Finally, the OECD Development Centre is investigating how Google data can help
elucidate governments’ approval in Latin America [@montoyaUsingGoogleData2020].

Research Question - How can we operationalize Google Search Trends as a valid
indicator for uses in social science research?

https://journals.sagepub.com/doi/10.1177/0894439316631043
			§ What constructs might google trends capture and not capture well? 
Capture attention but not attitudes

Also Asseo - "we assumed that media coverage may potentially decouple the search
popularity from the number of cases, since searches would result not only from
self-symptoms, but also from interest elicited by media coverage."

# Research Methodology
To investigate the construct and criterion validity of the use of Google Trends
in these three areas, I will gather geo-located data from a sources related to
each case and use Pearson’s r for pairwise correlations, testing for the
strength of relationship between the Google Trend indicator and the respective
comparison indicator. Pairwise correlations allow for testing of nested data,
allowing for multiple search terms to be compared across locations
simultaneously as well as comparing change over time. I will test both raw
questions with raw trends as well as using dimensionality reduction through
factor analysis and principal components analysis (PCA). In addition, I’d like
to explore how this data would cluster locations together using tools like
blockmodeling and other clustering techniques.

To investigate the correlation between Google search trends, I will use
questions from the NORC General Social Survey. I will need to obtain an $800
grant to purchase geolocated data from NORC for this project. I plan to submit a
small grant application to SBS to receive these funds. I will use various NORC
GSS cultural value questions and aggregate the relevant zip code up to Nielsen’s
metropolitan market area (MMA), the level Google trends are aggregated to, using
population weighting.

For disease prevalence, I will use United States county-level suicide rates from
National Center for Health Statistics, US cases rates of Covid-19 from The New
York Times, and obesity rates to compare search trends to observable data.

Finally, I’ll test Google Trends as an operationalization of political attitudes
by looking at actual voting outcomes in historical US Presidential Election
results and the American National Election Survey and Google Trends about
candidates.

```{r}
tibble::tribble(
                                ~Validated.source, ~Google.trend.equivalent,
                                            "GSS",                       NA,
  "Suicide,\\nCovid-19 rates\\nVaccine hesitancy",                       NA,
           "Presidential Voting outcomes,\\nANES",                       NA
  )

```

## Measures
### attitudinal
• Vaccine Hesitancy
		○ https://data.cdc.gov/Vaccinations/Vaccine-Hesitancy-for-COVID-19-County-and-local-es/q9mh-h2tw/data

• Mask Usage
		○ New York Times Survey

### health
• Covid Rates
		○ https://github.com/nytimes/covid-19-data/tree/master/rolling-averages

• Suicide
		○ https://wonder.cdc.gov/mcd.html 
		○ Multiple Cause of Death Data 2019 (see notes for which variables count)

### political
• Presidential results
		○ https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/VOQCHQ
		○ MIT Election Data and Science Lab, 2018, 
		○ "County Presidential Election Returns 2000-2020",
		○ https://doi.org/10.7910/DVN/VOQCHQ, 
		○ Harvard Dataverse, V9

### Independent Variables
## Analysis
# Results
## Cultural
## Medical
## Political
# Discussion
While I expect these tests to show high correlation between observed indicators
and Google search trends, there will be three important questions that surface.
First, just because something is correlated, does that mean it can replace the
collection of other types of data? Second, how correlated does a trend need to
be for social scientists to justifiably rely on it to indicate some outcome? And
finally, how can we construct analyses like this to be robust to changes in the
terms used across time and location?

The purpose of this paper is more methodological than theoretical, and I see
this paper having an impact on the social sciences and computational social
science as researchers pursue more projects using this source of big data.
Google Trends are relatively underutilized in the field compared to in the
health sciences and business. Once I assess how this data can be used, I would
like to be able to join @bailCulturalEnvironmentMeasuring2014 in encouraging social scientists to pursue
more research with big data while taking into account the potential pitfalls
with any source of big data [@mcfarlandBigDataDanger2015].

# Conclusion

\newpage

# References {-}

<div id="refs"></div>

\newpage

# Appendix