### 2022-06-18 paper1.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Article 1: Digital Trace Data as Indicators of Social Data Validating Google Trends for use in Scientific Research},
  pdfauthor={Kelsey Gonzalez},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{siunitx}
\usepackage{setspace}
\usepackage{dcolumn}
\usepackage{longtable}
\usepackage{caption}
\usepackage{booktabs}
\usepackage{placeins}
\usepackage{hhline}
\doublespacing
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi

\title{Article 1: Digital Trace Data as Indicators of Social Data
Validating Google Trends for use in Scientific Research}
\author{Kelsey Gonzalez}
\date{21 March, 2022}

\begin{document}
\maketitle

\hypertarget{research-methodology}{%
\section{Research Methodology}\label{research-methodology}}

\providecommand{\docline}[3]{\noalign{\global\setlength{\arrayrulewidth}{#1}}\arrayrulecolor[HTML]{#2}\cline{#3}}

\setlength{\tabcolsep}{2pt}

\renewcommand*{\arraystretch}{1.5}

\begin{longtable}[c]{cccc}

\caption{New York Air Quality Measurements
}(\#tab:data-sources-table)\\

\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}-}

\multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{\cellcolor[HTML]{004F63}}l}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{FFFFFF}{\global\setmainfont{Arial}{\textbf{Validated.Data.Source}}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{\cellcolor[HTML]{004F63}}l}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{FFFFFF}{\global\setmainfont{Arial}{\textbf{Type}}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{\cellcolor[HTML]{004F63}}l}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{FFFFFF}{\global\setmainfont{Arial}{\textbf{Dates}}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{\cellcolor[HTML]{004F63}}l!{\color[HTML]{666666}\vrule width 1pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{FFFFFF}{\global\setmainfont{Arial}{\textbf{Google.Trends.Used}}}}} \\

\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}-}

\endfirsthead

\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}-}

\multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{\cellcolor[HTML]{004F63}}l}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{FFFFFF}{\global\setmainfont{Arial}{\textbf{Validated.Data.Source}}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{\cellcolor[HTML]{004F63}}l}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{FFFFFF}{\global\setmainfont{Arial}{\textbf{Type}}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{\cellcolor[HTML]{004F63}}l}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{FFFFFF}{\global\setmainfont{Arial}{\textbf{Dates}}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{\cellcolor[HTML]{004F63}}l!{\color[HTML]{666666}\vrule width 1pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{FFFFFF}{\global\setmainfont{Arial}{\textbf{Google.Trends.Used}}}}} \\

\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}-}\endhead

\multicolumn{4}{!{\color[HTML]{666666}\vrule width 1pt}>{\cellcolor[HTML]{BAD7DB}}l!{\color[HTML]{666666}\vrule width 1pt}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{\textbf{\textit{Behaviors\ and\ Attitudes}}}}}} \\

\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}-}

\multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{General\ Social\ Survey}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{Cross-Sectional}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{2010\ -\ 2020}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l!{\color[HTML]{666666}\vrule width 1pt}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{}}}} \\

\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}-}

\multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{Vaccine\ Hesitancy\ for\ COVID-19}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{Cross-Sectional}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{March\ 3\ –\ 15,\ 2021}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l!{\color[HTML]{666666}\vrule width 1pt}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{Search\ Topics:\ 'Covid-19\ vaccine',\ 'Coronavirus\ (Disease)',\ 'Coronavirus\ (Virus)',\ 'Vaccine'}}}} \\

\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}-}

\multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{Mask-Wearing\ Survey\ Data}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{Cross-Sectional}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{July\ 2\ -\ 14,\ 2020}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l!{\color[HTML]{666666}\vrule width 1pt}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{Search\ Topics:\ 'Coronavirus\ (Disease)',\ 'Coronavirus\ (Virus)',\ 'Cloth\ Face\ Mask',\ 'Mask',\ 'Civil\ and\ Political\ Rights'}}}} \\

\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}-}

\multicolumn{4}{!{\color[HTML]{666666}\vrule width 1pt}>{\cellcolor[HTML]{BAD7DB}}l!{\color[HTML]{666666}\vrule width 1pt}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{\textbf{\textit{Health}}}}}} \\

\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}-}

\multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{Covid\ Rates}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{Longitudinal}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{Every\ Monday,\ 2020\ -\ 2021}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l!{\color[HTML]{666666}\vrule width 1pt}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{'Covid-19',\ 'Coronavirus',\ 'Taste\ Loss',\ 'Smell\ Loss'}}}} \\

\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}-}

\multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{County\ Suicide\ Rates}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{Longitudinal}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{Yearly\ 2010-2020}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l!{\color[HTML]{666666}\vrule width 1pt}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{Search\ Topics:\ 'Suicide',\ 'Depression',\ Search\ Term:'\ Suicide\ Hotline'}}}} \\

\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}-}

\multicolumn{4}{!{\color[HTML]{666666}\vrule width 1pt}>{\cellcolor[HTML]{BAD7DB}}l!{\color[HTML]{666666}\vrule width 1pt}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{\textbf{\textit{Political}}}}}} \\

\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}-}

\multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{American\ National\ Election\ Survey}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{Cross-Sectional}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{2020}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l!{\color[HTML]{666666}\vrule width 1pt}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{}}}} \\

\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}-}

\multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{Presidential\ Election\ Results}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{Cross\ Sectional}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{2016\ \&\ 2020}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l!{\color[HTML]{666666}\vrule width 1pt}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{Search\ Topics:\ 'Hilary\ Clinton',\ 'Donald\ Trump',\ 'Joe\ Biden'}}}} \\

\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}-}

\end{longtable}

\hypertarget{measures}{%
\subsection{Measures}\label{measures}}

\hypertarget{vacc_hes}{%
\paragraph{vacc\_hes}\label{vacc_hes}}

3148

from 4.99\% to 32.33\%.

\hypertarget{mask}{%
\paragraph{mask}\label{mask}}

(range = 0.10\% to\\
55.80\%)

\hypertarget{health}{%
\subsubsection{health}\label{health}}

\hypertarget{covid}{%
\paragraph{covid}\label{covid}}

0 to a maximum of 1460.46 for each Monday from January 27, 2020 through
December 27, 2021. There are 285986 cases across 3136 counties and 101
dates.

\hypertarget{suicide}{%
\paragraph{suicide}\label{suicide}}

There are 34683 total cases, resulting from 34617 observations of 3147
counties. Missing data were interpolated using.

Measures range from 0.034 to 53.254.

\hypertarget{political}{%
\subsubsection{political}\label{political}}

avaiable for 3147 counties in 2016\\
and 3118 counties in 2020.

with the lowest percent at 3.09\% and the highest at 92.15\%.

\providecommand{\docline}[3]{\noalign{\global\setlength{\arrayrulewidth}{#1}}\arrayrulecolor[HTML]{#2}\cline{#3}}

\setlength{\tabcolsep}{2pt}

\renewcommand*{\arraystretch}{1.5}

\begin{longtable}[c]{ccccccc}

\caption{Correlation Results
}(\#tab:corr-results)\\

\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}-}

\multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{\cellcolor[HTML]{004F63}}l}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{FFFFFF}{\global\setmainfont{Arial}{\textbf{measure}}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{\cellcolor[HTML]{004F63}}l}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{FFFFFF}{\global\setmainfont{Arial}{\textbf{variable}}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{\cellcolor[HTML]{004F63}}l}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{FFFFFF}{\global\setmainfont{Arial}{\textbf{trend1}}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{\cellcolor[HTML]{004F63}}l}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{FFFFFF}{\global\setmainfont{Arial}{\textbf{trend2}}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{\cellcolor[HTML]{004F63}}l}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{FFFFFF}{\global\setmainfont{Arial}{\textbf{trend3}}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{\cellcolor[HTML]{004F63}}l}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{FFFFFF}{\global\setmainfont{Arial}{\textbf{trend4}}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{\cellcolor[HTML]{004F63}}l!{\color[HTML]{666666}\vrule width 1pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{FFFFFF}{\global\setmainfont{Arial}{\textbf{trend5}}}}} \\

\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}-}

\endfirsthead

\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}-}

\multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{\cellcolor[HTML]{004F63}}l}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{FFFFFF}{\global\setmainfont{Arial}{\textbf{measure}}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{\cellcolor[HTML]{004F63}}l}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{FFFFFF}{\global\setmainfont{Arial}{\textbf{variable}}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{\cellcolor[HTML]{004F63}}l}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{FFFFFF}{\global\setmainfont{Arial}{\textbf{trend1}}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{\cellcolor[HTML]{004F63}}l}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{FFFFFF}{\global\setmainfont{Arial}{\textbf{trend2}}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{\cellcolor[HTML]{004F63}}l}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{FFFFFF}{\global\setmainfont{Arial}{\textbf{trend3}}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{\cellcolor[HTML]{004F63}}l}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{FFFFFF}{\global\setmainfont{Arial}{\textbf{trend4}}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{\cellcolor[HTML]{004F63}}l!{\color[HTML]{666666}\vrule width 1pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{FFFFFF}{\global\setmainfont{Arial}{\textbf{trend5}}}}} \\

\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}-}\endhead

\multicolumn{7}{!{\color[HTML]{666666}\vrule width 1pt}>{\cellcolor[HTML]{BAD7DB}}l!{\color[HTML]{666666}\vrule width 1pt}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{\textbf{\textit{Vaccine\ Hesitancy}}}}}} \\

\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}-}

\multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{Covid-19\ vaccine}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{Coronavirus\ (Disease)}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{Coronavirus\ (Virus)}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{Vaccine}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l!{\color[HTML]{666666}\vrule width 1pt}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{}}}} \\

\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}-}

\multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{Pearson's\ R\ Correlation}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{Vaccine\ Hesitancy}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{-0.3866}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{-0.4657}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{-0.3561}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{-0.3994}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l!{\color[HTML]{666666}\vrule width 1pt}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{}}}} \\

\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}-}

\multicolumn{7}{!{\color[HTML]{666666}\vrule width 1pt}>{\cellcolor[HTML]{BAD7DB}}l!{\color[HTML]{666666}\vrule width 1pt}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{\textbf{\textit{Mask\ Attitudes}}}}}} \\

\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}-}

\multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{Coronavirus\ (Disease)}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{Coronavirus\ (Virus)}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{Cloth\ Face\ Mask}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{Mask}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l!{\color[HTML]{666666}\vrule width 1pt}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{Civil\ and\ Political\ Rights}}}} \\

\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}-}

\multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{Pearson's\ R\ Correlation}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{Mask\ Rare}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{-0.1398}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{-0.1343}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{0.146}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{0.1339}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l!{\color[HTML]{666666}\vrule width 1pt}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{-0.1175}}}} \\

\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}-}

\multicolumn{7}{!{\color[HTML]{666666}\vrule width 1pt}>{\cellcolor[HTML]{BAD7DB}}l!{\color[HTML]{666666}\vrule width 1pt}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{\textbf{\textit{Covid\ Rates}}}}}} \\

\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}-}

\multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{covid\_19}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{smell\_loss}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{taste\_loss}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l!{\color[HTML]{666666}\vrule width 1pt}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{}}}} \\

\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}-}

\multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{Pearson's\ R\ Correlation}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{Covid\ Rate}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{-0.1012}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{0.2835}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{0.2851}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l!{\color[HTML]{666666}\vrule width 1pt}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{}}}} \\

\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}-}

\multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{repeated\ measures\ correlation\ coefficient}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{Covid\ Rate}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{-0.099***}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{0.322***}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{0.313***}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l!{\color[HTML]{666666}\vrule width 1pt}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{}}}} \\

\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}-}

\multicolumn{7}{!{\color[HTML]{666666}\vrule width 1pt}>{\cellcolor[HTML]{BAD7DB}}l!{\color[HTML]{666666}\vrule width 1pt}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{\textbf{\textit{Suicide\ Rates}}}}}} \\

\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}-}

\multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{suicide}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{depression}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{suicide\_hotline}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l!{\color[HTML]{666666}\vrule width 1pt}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{}}}} \\

\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}-}

\multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{Pearson's\ R\ Correlation}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{Suicide\ Rate}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{0.0845}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{0.0273}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{0.0618}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l!{\color[HTML]{666666}\vrule width 1pt}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{}}}} \\

\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}-}

\multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{repeated\ measures\ correlation\ coefficient}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{Suicide\ Rate}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{0.055***}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{0.142***}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{0.131***}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l!{\color[HTML]{666666}\vrule width 1pt}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{}}}} \\

\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}-}

\multicolumn{7}{!{\color[HTML]{666666}\vrule width 1pt}>{\cellcolor[HTML]{BAD7DB}}l!{\color[HTML]{666666}\vrule width 1pt}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{\textbf{\textit{2016\ Presidential\ Votes}}}}}} \\

\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}-}

\multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{Hilary\ Clinton}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{Donald\ Trump}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l!{\color[HTML]{666666}\vrule width 1pt}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{}}}} \\

\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}-}

\multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{Pearson's\ R\ Correlation}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{2016\ Votes\ for\ Clinton}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{-0.1663}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{0.1663}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l!{\color[HTML]{666666}\vrule width 1pt}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{}}}} \\

\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}-}

\multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{Pearson's\ R\ Correlation}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{2016\ Votes\ for\ Trump}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{0.1682}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{-0.1682}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l!{\color[HTML]{666666}\vrule width 1pt}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{}}}} \\

\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}-}

\multicolumn{7}{!{\color[HTML]{666666}\vrule width 1pt}>{\cellcolor[HTML]{BAD7DB}}l!{\color[HTML]{666666}\vrule width 1pt}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{\textbf{\textit{2020\ Presidential\ Votes}}}}}} \\

\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}-}

\multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{Joe\ Biden}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{Donald\ Trump}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l!{\color[HTML]{666666}\vrule width 1pt}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{}}}} \\

\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}-}

\multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{Pearson's\ R\ Correlation}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{2020\ Votes\ for\ Biden}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{-0.0286}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{0.0286}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l!{\color[HTML]{666666}\vrule width 1pt}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{}}}} \\

\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}-}

\multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{Pearson's\ R\ Correlation}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{2020\ Votes\ for\ Trump}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{0.0356}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{-0.0356}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{}l!{\color[HTML]{666666}\vrule width 1pt}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\global\setmainfont{Arial}{}}}} \\

\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}-}

\end{longtable}

\begin{center}\includegraphics[width=0.8\linewidth]{paper1_files/figure-latex/vacc_hes_plot-1} \end{center}

\begin{table}

\caption{\label{tab:vacc_hes_analysis}Linear Regression Results for Vaccine Hesitancy}
\centering
\begin{tabular}[t]{lccc}
\toprule
  & Model 1 & Model 2 & Model 3\\
\midrule
(Intercept) & \num{0.191}*** & \num{0.191}*** & \num{0.191}***\\
 & (\num{0.001}) & (\num{0.001}) & (\num{0.001})\\
covid\_19\_vaccine & \num{-0.006}+ &  & \num{-0.011}**\\
 & (\num{0.004}) &  & (\num{0.003})\\
vaccine & \num{0.021}*** &  & \num{0.023}***\\
 & (\num{0.005}) &  & (\num{0.005})\\
coronavirus\_disease & \num{-0.037}*** &  & \num{-0.027}***\\
 & (\num{0.003}) &  & (\num{0.003})\\
coronavirus\_virus & \num{-0.003}* &  & \num{0.000}\\
 & (\num{0.001}) &  & (\num{0.001})\\
total\_pop &  & \num{-0.005}*** & \num{-0.004}***\\
 &  & (\num{0.001}) & \vphantom{5} (\num{0.001})\\
pop\_density &  & \num{-0.004}*** & \num{-0.003}***\\
 &  & (\num{0.001}) & \vphantom{4} (\num{0.001})\\
unemployment\_rate &  & \num{-0.002}* & \num{0.000}\\
 &  & (\num{0.001}) & \vphantom{3} (\num{0.001})\\
over\_65 &  & \num{-0.013}*** & \num{-0.010}***\\
 &  & (\num{0.001}) & \vphantom{2} (\num{0.001})\\
poverty\_rate &  & \num{0.005}** & \num{0.004}**\\
 &  & (\num{0.002}) & (\num{0.001})\\
median\_income &  & \num{-0.021}*** & \num{-0.016}***\\
 &  & (\num{0.001}) & \vphantom{1} (\num{0.001})\\
broadband &  & \num{-0.008}*** & \num{-0.005}***\\
 &  & (\num{0.001}) & (\num{0.001})\\
\midrule
Num.Obs. & \num{3148} & \num{3133} & \num{3133}\\
R2 & \num{0.233} & \num{0.350} & \num{0.411}\\
R2 Adj. & \num{0.232} & \num{0.349} & \num{0.409}\\
Log.Lik. & \num{5163.343} & \num{5405.127} & \num{5557.567}\\
F & \num{238.587} & \num{240.876} & \num{197.733}\\
\bottomrule
\multicolumn{4}{l}{\rule{0pt}{1em}* p $<$ .05. ** p $<$ .01. *** p $<$ .001 (two-tailed test).}\\
\end{tabular}
\end{table}

\begin{center}\includegraphics[width=0.8\linewidth]{paper1_files/figure-latex/mask_plot-1} \end{center}

\begin{table}

\caption{\label{tab:mask_analysis}Linear Regression Results for Rare Mask Usage}
\centering
\begin{tabular}[t]{lccc}
\toprule
  & Model 1 & Model 2 & Model 3\\
\midrule
(Intercept) & \num{0.162}*** & \num{0.163}*** & \num{0.162}***\\
 & (\num{0.002}) & (\num{0.002}) & (\num{0.002})\\
coronavirus\_disease & \num{-0.011}*** &  & \num{-0.012}***\\
 & (\num{0.003}) &  & \vphantom{3} (\num{0.003})\\
coronavirus\_virus & \num{-0.007}* &  & \num{0.003}\\
 & (\num{0.003}) &  & \vphantom{2} (\num{0.003})\\
cloth\_face\_mask & \num{-0.003} &  & \num{-0.008}*\\
 & (\num{0.003}) &  & \vphantom{1} (\num{0.003})\\
mask & \num{0.019}*** &  & \num{0.016}***\\
 & (\num{0.003}) &  & (\num{0.003})\\
total\_pop &  & \num{-0.012}*** & \num{-0.011}***\\
 &  & (\num{0.002}) & \vphantom{3} (\num{0.002})\\
pop\_density &  & \num{-0.003}+ & \num{-0.003}*\\
 &  & (\num{0.002}) & \vphantom{2} (\num{0.002})\\
unemployment\_rate &  & \num{-0.028}*** & \num{-0.027}***\\
 &  & (\num{0.002}) & \vphantom{1} (\num{0.002})\\
over\_65 &  & \num{-0.006}** & \num{-0.005}*\\
 &  & (\num{0.002}) & (\num{0.002})\\
poverty\_rate &  & \num{-0.010}** & \num{-0.010}**\\
 &  & (\num{0.003}) & \vphantom{2} (\num{0.003})\\
median\_income &  & \num{-0.036}*** & \num{-0.033}***\\
 &  & (\num{0.003}) & \vphantom{1} (\num{0.003})\\
broadband &  & \num{-0.006}* & \num{-0.006}*\\
 &  & (\num{0.003}) & (\num{0.003})\\
\midrule
Num.Obs. & \num{3136} & \num{3133} & \num{3121}\\
R2 & \num{0.049} & \num{0.176} & \num{0.189}\\
R2 Adj. & \num{0.047} & \num{0.174} & \num{0.186}\\
Log.Lik. & \num{2970.506} & \num{3184.261} & \num{3202.189}\\
F & \num{40.072} & \num{95.547} & \num{65.970}\\
\bottomrule
\multicolumn{4}{l}{\rule{0pt}{1em}* p $<$ .05. ** p $<$ .01. *** p $<$ .001 (two-tailed test).}\\
\end{tabular}
\end{table}

\begin{center}\includegraphics[width=0.8\linewidth]{paper1_files/figure-latex/covid_plot-1} \end{center}

Random effect variances not available. Returned R2 does not account for
random effects. Random effect variances not available. Returned R2 does
not account for random effects.

\begin{table}

\caption{\label{tab:covid_analysis}Hierarchical Model for Covid Case Rates}
\centering
\begin{tabular}[t]{lccc}
\toprule
  & Model 1 & Model 2 & Model 3\\
\midrule
(Intercept) & \num{-532.306}*** & \num{-712.023}*** & \num{-534.608}***\\
 & (\num{8.271}) & (\num{6.009}) & (\num{8.268})\\
covid\_19 & \num{1.031}*** &  & \num{1.024}***\\
 & (\num{0.082}) &  & \vphantom{1} (\num{0.082})\\
smell\_loss & \num{7.373}*** &  & \num{7.374}***\\
 & (\num{0.082}) &  & (\num{0.082})\\
taste\_loss & \num{6.344}*** &  & \num{6.347}***\\
 & (\num{0.078}) &  & (\num{0.078})\\
covid\_rate\_fips\_mean & \num{0.813}*** & \num{0.982}*** & \num{0.879}***\\
 & (\num{0.019}) & (\num{0.010}) & (\num{0.020})\\
date & \num{0.029}*** & \num{0.038}*** & \num{0.029}***\\
 & (\num{0.000}) & (\num{0.000}) & (\num{0.000})\\
SD (Intercept) & \num{5.196} & \num{0.002} & \num{4.984}\\
SD (Observations) & \num{29.478} & \num{31.736} & \num{29.441}\\
total\_pop &  & \num{0.064} & \num{0.409}**\\
 &  & (\num{0.066}) & (\num{0.140})\\
pop\_density &  & \num{0.010} & \num{0.463}***\\
 &  & (\num{0.067}) & (\num{0.123})\\
unemployment\_rate &  & \num{0.049} & \num{0.725}***\\
 &  & (\num{0.080}) & (\num{0.155})\\
over\_65 &  & \num{-0.116} & \num{-0.055}\\
 &  & (\num{0.074}) & (\num{0.139})\\
poverty\_rate &  & \num{0.054} & \num{0.351}\\
 &  & (\num{0.118}) & (\num{0.221})\\
median\_income &  & \num{0.031} & \num{0.907}***\\
 &  & (\num{0.114}) & (\num{0.214})\\
broadband &  & \num{0.073} & \num{0.437}*\\
 &  & (\num{0.097}) & (\num{0.182})\\
\midrule
Num.Obs. & \num{248346} & \num{284666} & \num{247655}\\
R2 Marg. & \num{0.201} & \num{0.081} & \num{0.200}\\
R2 Cond. &  & \num{0.081} & \\
\bottomrule
\multicolumn{4}{l}{\rule{0pt}{1em}Random intercept per county}\\
\multicolumn{4}{l}{\rule{0pt}{1em}* p $<$ .05. ** p $<$ .01. *** p $<$ .001 (two-tailed test).}\\
\end{tabular}
\end{table}

\begin{center}\includegraphics[width=0.8\linewidth]{paper1_files/figure-latex/suicide_plot-1} \end{center}

Random effect variances not available. Returned R2 does not account for
random effects. Random effect variances not available. Returned R2 does
not account for random effects. Random effect variances not available.
Returned R2 does not account for random effects.

\begin{table}

\caption{\label{tab:suicide_analysis}Hierarchical Model for Suicide Rates}
\centering
\begin{tabular}[t]{lccc}
\toprule
  & Model 1 & Model 2 & Model 3\\
\midrule
(Intercept) & \num{-7.364}*** & \num{-10.046}*** & \num{-9.683}***\\
 & (\num{0.359}) & (\num{0.342}) & (\num{0.454})\\
suicide & \num{-0.001} &  & \num{-0.001}\\
 & (\num{0.001}) &  & \vphantom{2} (\num{0.001})\\
depression & \num{-0.001}+ &  & \num{-0.001}\\
 & (\num{0.001}) &  & \vphantom{1} (\num{0.001})\\
suicide\_hotline & \num{0.002}*** &  & \num{0.002}***\\
 & (\num{0.001}) &  & (\num{0.001})\\
year & \num{0.004}*** & \num{0.005}*** & \num{0.005}***\\
 & (\num{0.000}) & (\num{0.000}) & (\num{0.000})\\
SD (Intercept) & \num{1.460} & \num{1.587} & \num{1.459}\\
SD (Observations) & \num{0.050} & \num{0.051} & \num{0.050}\\
total\_pop &  & \num{-0.032}*** & \num{-0.027}**\\
 &  & (\num{0.010}) & (\num{0.009})\\
pop\_density &  & \num{-0.037}** & \num{-0.033}**\\
 &  & (\num{0.011}) & (\num{0.011})\\
unemployment\_rate &  & \num{-0.001} & \num{-0.001}\\
 &  & (\num{0.001}) & \vphantom{1} (\num{0.001})\\
over\_65 &  & \num{-0.010}*** & \num{-0.014}***\\
 &  & (\num{0.002}) & \vphantom{1} (\num{0.002})\\
poverty\_rate &  & \num{-0.005}*** & \num{-0.003}**\\
 &  & (\num{0.001}) & (\num{0.001})\\
median\_income &  & \num{-0.014}*** & \num{-0.015}***\\
 &  & (\num{0.002}) & (\num{0.002})\\
\midrule
Num.Obs. & \num{30008} & \num{33582} & \num{30001}\\
R2 Marg. & \num{0.058} & \num{0.582} & \num{0.564}\\
R2 Cond. &  &  & \\
\bottomrule
\multicolumn{4}{l}{\rule{0pt}{1em}Random intercept per county}\\
\multicolumn{4}{l}{\rule{0pt}{1em}* p $<$ .05. ** p $<$ .01. *** p $<$ .001 (two-tailed test).}\\
\end{tabular}
\end{table}

\begin{center}\includegraphics[width=0.8\linewidth]{paper1_files/figure-latex/pres_plot-1} \end{center}

\begin{table}

\caption{\label{tab:pres_2016_analysis}Linear Regression Results for 2016 Presidential Election Results (Hilary Clinton Shown)}
\centering
\begin{tabular}[t]{lccc}
\toprule
  & Model 1 & Model 2 & Model 3\\
\midrule
(Intercept) & \num{0.318}*** & \num{0.318}*** & \num{0.318}***\\
 & (\num{0.003}) & (\num{0.002}) & (\num{0.002})\\
hillary\_clinton\_trend & \num{-0.025}*** &  & \num{-0.019}***\\
 & (\num{0.003}) &  & (\num{0.002})\\
total\_pop &  & \num{0.028}*** & \num{0.028}***\\
 &  & (\num{0.002}) & \vphantom{1} (\num{0.002})\\
pop\_density &  & \num{0.022}*** & \num{0.022}***\\
 &  & (\num{0.002}) & (\num{0.002})\\
unemployment\_rate &  & \num{0.047}*** & \num{0.045}***\\
 &  & (\num{0.003}) & \vphantom{1} (\num{0.003})\\
over\_65 &  & \num{-0.012}*** & \num{-0.014}***\\
 &  & (\num{0.003}) & (\num{0.003})\\
poverty\_rate &  & \num{0.034}*** & \num{0.034}***\\
 &  & (\num{0.004}) & \vphantom{1} (\num{0.004})\\
median\_income &  & \num{0.054}*** & \num{0.052}***\\
 &  & (\num{0.004}) & (\num{0.004})\\
\midrule
Num.Obs. & \num{3147} & \num{3145} & \num{3145}\\
R2 & \num{0.028} & \num{0.314} & \num{0.329}\\
R2 Adj. & \num{0.027} & \num{0.313} & \num{0.327}\\
Log.Lik. & \num{1483.493} & \num{2037.065} & \num{2071.025}\\
\bottomrule
\multicolumn{4}{l}{\rule{0pt}{1em}Results predicting Donald J. Trump percentage largely equivalent and available upon request.}\\
\multicolumn{4}{l}{\rule{0pt}{1em}* p $<$ .05. ** p $<$ .01. *** p $<$ .001 (two-tailed test).}\\
\end{tabular}
\end{table}

\begin{table}

\caption{\label{tab:pres_2020_analysis}Linear Regression Results for 2020 Presidential Election Results (Joe Biden Shown)}
\centering
\begin{tabular}[t]{lccc}
\toprule
  & Model 1 & Model 2 & Model 3\\
\midrule
(Intercept) & \num{0.333}*** & \num{0.333}*** & \num{0.333}***\\
 & (\num{0.003}) & (\num{0.002}) & (\num{0.002})\\
joe\_biden\_trend & \num{-0.005} &  & \num{-0.014}***\\
 & (\num{0.003}) &  & (\num{0.002})\\
total\_pop &  & \num{0.030}*** & \num{0.029}***\\
 &  & (\num{0.003}) & \vphantom{3} (\num{0.003})\\
pop\_density &  & \num{0.021}*** & \num{0.022}***\\
 &  & (\num{0.003}) & \vphantom{2} (\num{0.003})\\
unemployment\_rate &  & \num{0.044}*** & \num{0.044}***\\
 &  & (\num{0.003}) & \vphantom{1} (\num{0.003})\\
over\_65 &  & \num{-0.007}* & \num{-0.008}**\\
 &  & (\num{0.003}) & (\num{0.003})\\
poverty\_rate &  & \num{0.032}*** & \num{0.033}***\\
 &  & (\num{0.004}) & \vphantom{1} (\num{0.004})\\
median\_income &  & \num{0.069}*** & \num{0.070}***\\
 &  & (\num{0.004}) & (\num{0.004})\\
\midrule
Num.Obs. & \num{3118} & \num{3118} & \num{3118}\\
R2 & \num{0.001} & \num{0.295} & \num{0.303}\\
R2 Adj. & \num{0.000} & \num{0.294} & \num{0.301}\\
Log.Lik. & \num{1291.161} & \num{1835.433} & \num{1852.002}\\
\bottomrule
\multicolumn{4}{l}{\rule{0pt}{1em}Results predicting Donald J. Trump percentage largely equivalent and available upon request.}\\
\multicolumn{4}{l}{\rule{0pt}{1em}* p $<$ .05. ** p $<$ .01. *** p $<$ .001 (two-tailed test).}\\
\end{tabular}
\end{table}

\end{document}